---
title: "PSTAT220A Homework 4"
author: "sbsambado"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(faraway)
```

## Question 1
```{r}
# upload data
teen <- teengamb

# run linear model
mod1 <- lm(gamble ~ income, data = teen)

# plot model results
plot(mod1)
```

#### A) 
Based on the 'Residuals vs Fitted' graph, there is a slight pattern but because most residuals fall along the 0 line and the pattern is not cone shaped I think mod1 (income ~ income) does not violate the constant variance assumption for the purposes of this class. 

#### B)
Based on the 'Normal Q-Q' graph, majority of the points fall along the 1:1 line so I think mod1 meets the normality assumptions. Although there are points such as 17, 36, 24 that are some distance from the 1:1 line, I still would deem this data normally distributed until I further investigate how these outliers were collected and their biological relevance.

#### C)
Based on the 'Scale-Location' graph, there are large leverage points such as 36,24,17 and occur when income is relatively large. 

#### D)
Based on the 'Relevance vs Leverage'

